---
title: "LightLogR for metric calculation"
format: 
  html:
    self-contained: true
    code-link: true
    code-tools: true
---

## Preface

This script summarizes simulation runs of light exposure in terms of light exposure metrics and time-based values, both for tables and plots. The relevant data are stored in files that conform to `results/.*_sequence.csv`. Outputs are stored in `results/plots/LightLogR` and `results/tables/LightLogR`.

## Setup

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(LightLogR)
library(gt)
library(glue)
library(ggtext)
library(legendry)
library(here)
library(rnaturalearth)
library(rnaturalearthdata)
library(melidosData)
library(ggrepel)
library(patchwork)
tz <- "America/Toronto"
coordinates <- c(45.449, -75.625)
```

## Worldmap

```{r}
# source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/overview_plot.R")
```

```{r}
colors_plot <- c(
melidos_colors[c("IZTECH", "FUSPCEU", "MPI", "KNUST", "UCR")],
ML = "#0073C2FF",
CH = "#EFC000FF",
CA = "#D80621",
UK = "blue"
)

countries <- c(
  melidos_countries[c("IZTECH", "FUSPCEU", "MPI", "KNUST", "UCR")],
  ML = "Malaysia",
  CH = "Switzerland",
  CA = "Canada",
  UK = "United Kindgom"
  )

locations <- c(
  melidos_cities[c("IZTECH", "FUSPCEU", "MPI", "KNUST", "UCR")],
  ML = "Kuala Lumpur",
  CH = "Basel",
  CA = "Ottawa",
  UK = "Manchester"
)

lats <- c(
  melidos_coordinates[c("IZTECH", "FUSPCEU", "MPI", "KNUST", "UCR")] |> map(1) |> unlist(),
  ML = 3.0650,
  CH = 47.5585,
  CA = coordinates[1],
  UK = 53.47
)
lons <- c(
  melidos_coordinates[c("IZTECH", "FUSPCEU", "MPI", "KNUST", "UCR")] |> map(2) |> unlist(),
  ML = 101.6009,
  CH = 7.5839,
  CA = coordinates[2],
  UK = -2.23
)
```


```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
    
    countries_colors <- tibble(
      country = countries,
      color   = colors_plot,
      stringsAsFactors = FALSE
    )
    
    world$color <- ifelse(
      world$name %in% countries_colors$country,
      countries_colors$country[match(world$name, countries_colors$country)],
      NA
    )
    
    # 3) Locations in geographic CRS (EPSG:4326)
    location_info <- tibble(
      country  = countries,
      location = locations,
      lat      = lats,
      lon      = lons,
      color    = colors_plot,
      stringsAsFactors = FALSE
    ) |> 
      rowwise() |> 
      mutate(
        coordinate_string = format_coordinates(c(lat,lon)),
        label = 
          paste0(location, ", ", country, " (", coordinate_string, ")")
        )
    
    locations <- st_as_sf(location_info, coords = c("lon", "lat"), crs = 4326)
    
    # 4) Project both layers to a planar CRS
    # robinson_crs <- paste0("+proj=", map_projection)
    robinson_crs <- "+proj=eqc"
    world_proj     <- st_transform(world,    crs = robinson_crs)
    locations_proj <- st_transform(locations, crs = robinson_crs)
    
    bb <- st_bbox(world_proj)
    y_offset <- 0.08 * (bb["ymax"] - bb["ymin"]) 
    
    # 5) Plot
```


```{r}
#| fig-width: 10
#| fig-height: 5
ggplot() +
      geom_sf(data = world_proj, aes(fill = color),
              color = NA, size = 0.25, alpha = 0.5, show.legend = FALSE) +
      geom_sf(data = locations_proj, aes(fill = country),
              shape = 21, color = "black", size = 3, stroke = 0.2) +
      geom_label_repel(
        data = locations_proj,
        stat = "sf_coordinates",
        label.size = 0,
        aes(label = label, fill = country, geometry = geometry),
        color = c(rep("black", 7),  "white", "white"),
        min.segment.length = 0.75,
        size = 3,
        alpha = 0.8,
        box.padding = 0.5,    # space between labels
        max.overlaps = 20,
        point.padding = 0.25,  # space from point to label
        force = 5,             # stronger repulsion
        force_pull = 0.25,      # weaker pull back to point
        seed = 123
      ) +
      scale_fill_manual(values = 
                          locations_proj |> 
                          as_tibble() |> 
                          select(country, color) |> 
                          deframe()) +
      theme_cowplot() +
      theme(legend.position = "none") +
      theme_sub_axis(line = element_blank()) +
      labs(x = NULL, y = NULL) +
      coord_sf(expand = FALSE, clip = "off") +
  theme_sub_plot(margin = margin(5, 12, 5, 5))
```


```{r}
ggsave(here("results/plots/LightLogR/worldmap.pdf"), width = 10, height = 5)
```

```{r}
#| label: worldmap
#| eval:false
grand_overview("", coordinates, "Ottawa", "Canada", "#D80621", panels = "map")
ggsave("../results/plots/LightLogR/worldmap.pdf", height = 5, width = 10)
```

## Import

```{r}
#| label: import
path <- "../results/tables"
files <- list.files(path, pattern = "_sequence.csv", full.names = TRUE)
data <- files |> set_names() |> map(read_csv, .progress = "Loading in data")
data <-
  data |>
  map({
    \(x) x |>
      rename(Datetime = "...1") |>
      mutate(Datetime =
               lubridate::ymd_hms(
                 glue(
                   "2023-06-16 {
                  hms::hms(minutes = seq(0, 30*21, by = 30), hours = rep(8, 22))
             }"
                 ),tz = tz
               )) |>
      pivot_longer(
        -Datetime,
        names_to = "Sequence",
        names_prefix = "seq_",
        values_to = "MEDI"
      )
  }, .progress = "Preparing loaded data")

names(data) <- names(data) |> basename() |> str_remove("_sequence\\.csv")
```

## Metrics

Metrics to calculate, all based on the calculated values for melanopic EDI (lx):

- $TAT_{250}$ (Time above 250, lx)
- $TAT_{1000}$ (Time above 1000, lx)
- $Dose$ (Dose, lx·h)
- $MLiT_{250}$ (Mean timing of light above 250, HH:MM)
- $FLiT_{250}$ (First timing of light above 250, HH:MM)

```{r}
#| label: calculate metrics
metrics_raw <-
  data |>
  map({
    \(x) x |>
  group_by(Sequence) |> 
  summarize(
    dose(MEDI, Datetime, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 250, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 1000, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 10000, as.df = TRUE),
  )
  }, .progress = "Calculating metrics")

metrics <-
  metrics_raw |>
  map({
    \(x) x |>
  summarize(
    across(dose:duration_above_10000,
           .fns = list(mean = ~mean(.x, na.rm = TRUE),
                       median = ~median(.x, na.rm = TRUE),
                       sd = ~sd(.x, na.rm = TRUE),
                       min = ~min(.x, na.rm = TRUE),
                       max = ~max(.x, na.rm = TRUE),
                       q25 = ~quantile(.x, 0.25, na.rm = TRUE),
                       q75 = ~quantile(.x, 0.75, na.rm = TRUE)
                       ),
           .names = "{.col}.{.fn}"
           )
    ) |>
  pivot_longer(
    cols = everything(),
    names_to   = c("metric", "stat"),
    names_sep  = "\\.",
    values_transform = list(value = as.numeric)  # mix of dbl/duration/POSIXct
  ) |>
  pivot_wider(
    names_from  = metric,        # columns = metrics
    values_from = value          # rows = statistics
  ) |>
  mutate(
    across(-stat, round),
    across(contains("duration"), as.duration),
    across(contains("timing"), hms::as_hms)
    )
  }, .progress = "Calculating metrics")

metrics |> 
  imap(\(x, idx) x |> 
                  gt() |> 
                  cols_label_with(fn = \(x) x |> 
                               str_replace_all("_", " ") |> 
                               str_to_sentence()) |> 
                  tab_header(idx |> 
                              str_replace_all("_", " ") |> 
                               str_to_sentence()
                             )
                  
       )

metrics_raw |> 
  iwalk(\(x, idx) {
       write_csv(x, glue("../results/tables/LightLogR/metrics_raw/{idx}.csv"))
       write_csv(x |> mutate(across(-Sequence, as.numeric)), 
                 glue("../results/tables/LightLogR/metrics_raw/{idx}_numeric.csv"))
    },
       .progress = "Writing metrics")

metrics |> 
  iwalk(\(x, idx) {
       write_csv(x, glue("../results/tables/LightLogR/metrics/{idx}.csv"))
       write_csv(x |> mutate(across(-stat, as.numeric)), 
                 glue("../results/tables/LightLogR/metrics/{idx}_numeric.csv"))
    },
       .progress = "Writing metrics")
```

## Timeline

### Data

Stats for each time slot from 8:00 until 18:30.

```{r}
#| label: calculate timeline stats
time_stats <-
  data |>
  map({
    \(x) x |>
  # filter(Sequence %in% c("0001", "0002", "0003", "0004")) |>
      select(-Sequence) |> 
  aggregate_Datetime(
    unit = "30 mins",
    mean = mean(MEDI, na.rm = TRUE),
    median = median(MEDI, na.rm = TRUE),
    sd = sd(MEDI, na.rm = TRUE),
    min = min(MEDI, na.rm = TRUE),
    max = max(MEDI, na.rm = TRUE),
    q25 = quantile(MEDI, 0.25, na.rm = TRUE),
    q75 = quantile(MEDI, 0.75, na.rm = TRUE),
    n. = n()
    ) |> 
    select(-MEDI) |> 
      rename(n = n.)
  }, .progress = "Calculating Timestats")

time_stats |> 
  imap(\(x, idx) x |> 
                  gt() |> 
                  cols_label_with(fn = \(x) x |> 
                               str_replace_all("_", " ") |> 
                               str_to_sentence()) |> 
                  tab_header(idx |> 
                              str_replace_all("_", " ") |> 
                               str_to_sentence()
                             )
                  
       )

time_stats |> 
  iwalk(\(x, idx) {
       write_csv(x, glue("../results/tables/LightLogR/timeline/{idx}.csv"))
    },
       .progress = "Writing timeline")

```

### Plot

```{r}
#| label: create plots
caption <- glue("<strong>mean</strong>, <strong style = 'color:{alpha('#0073C2FF', 
          alpha = 1)}'>IQR</strong>, <strong style = 'color:{alpha('#0073C2FF', 
          alpha = 0.25)}'>range</strong>")

time_stats_plot <-
  time_stats |>
  imap({
   \(x, idx) x |> 
  gg_day(mean, geom = "blank",
         facetting = FALSE,
         y.axis.label = "melanopic EDI (lx)"
         ) |> 
  gg_photoperiod(coordinates) +
  geom_ribbon(aes(ymin = min, ymax = max), fill = "#0073C2FF", alpha = 0.25) +
  geom_ribbon(aes(ymin = q25, ymax = q75), fill = "#0073C2FF") +
  geom_line() +
  scale_y_continuous(trans = "symlog", 
                     labels = expression(0,10^0,10^1, 10^2, 10^3, 10^4, 10^5),
                     breaks = c(0, 10^(0:5))) +
  geom_point() +
      coord_cartesian(ylim = c(1,10^5)) +
      labs(caption = caption, title = idx |> 
                                        str_replace_all("_", " ") |> 
                                        str_to_sentence()) +
      theme_sub_plot(caption = element_textbox_simple())
  })

time_stats_plot

time_stats_plot |> 
  iwalk(\(x, idx) {
       ggsave(plot = x, glue("../results/plots/LightLogR/Timeline_{idx}.pdf"),
              height = 4, width = 6)
    },
       .progress = "Saving timeline plots")

```

