---
title: "LightLogR for metric calculation"
format: 
  html:
    self-contained: true
    code-link: true
    code-tools: true
---

## Preface

This script summarizes simulation runs of light exposure in terms of light exposure metrics and time-based values, both for tables and plots. The relevant data are stored in files that conform to `results/.*_sequence.csv`. Outputs are stored in `results/plots/LightLogR` and `results/tables/LightLogR`.

## Setup

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(LightLogR)
library(gt)
library(glue)
library(ggtext)
library(legendry)
tz <- "America/Toronto"
coordinates <- c(45.449, -75.625)
```

## Worldmap

```{r}
#| label: worldmap
#| eval:false

source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/main/scripts/overview_plot.R")

grand_overview("", coordinates, "Ottawa", "Canada", "#D80621", panels = "map")
ggsave("../results/plots/LightLogR/worldmap.pdf", height = 3.4)
```

## Import

```{r}
#| label: import
path <- "../results"
files <- list.files(path, pattern = "_sequence.csv", full.names = TRUE)
data <- files |> set_names() |> map(read_csv, .progress = "Loading in data")
data <-
  data |>
  map({
    \(x) x |>
      rename(Datetime = "...1") |>
      mutate(Datetime =
               lubridate::ymd_hms(
                 glue(
                   "2023-06-16 {
                  hms::hms(minutes = seq(0, 30*21, by = 30), hours = rep(8, 22))
             }"
                 ),tz = tz
               )) |>
      pivot_longer(
        -Datetime,
        names_to = "Sequence",
        names_prefix = "seq_",
        values_to = "MEDI"
      )
  }, .progress = "Preparing loaded data")

names(data) <- names(data) |> basename() |> str_remove("_sequence\\.csv")
```

## Metrics

Metrics to calculate, all based on the calculated values for melanopic EDI (lx):

- $TAT_{250}$ (Time above 250, lx)
- $TAT_{1000}$ (Time above 1000, lx)
- $Dose$ (Dose, lx·h)
- $MLiT_{250}$ (Mean timing of light above 250, HH:MM)
- $FLiT_{250}$ (First timing of light above 250, HH:MM)

```{r}
#| label: calculate metrics
metrics <-
  data |>
  map({
    \(x) x |>
  # filter(Sequence %in% c("0001", "0002", "0003", "0004")) |> 
  group_by(Sequence) |> 
  summarize(
    dose(MEDI, Datetime, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 250, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 1000, as.df = TRUE),
    timing_above_threshold(MEDI, Datetime, threshold = 250, as.df = TRUE)
  ) |> 
  Datetime2Time() |> 
  summarize(
    across(dose:last_timing_above_250,
           .fns = list(mean = ~mean(.x, na.rm = TRUE),
                       median = ~median(.x, na.rm = TRUE),
                       sd = ~sd(.x, na.rm = TRUE),
                       min = ~min(.x, na.rm = TRUE),
                       max = ~max(.x, na.rm = TRUE),
                       q25 = ~quantile(.x, 0.25, na.rm = TRUE),
                       q75 = ~quantile(.x, 0.75, na.rm = TRUE)
                       ),
           .names = "{.col}.{.fn}"
           )
    ) |> 
  pivot_longer(
    cols = everything(),
    names_to   = c("metric", "stat"),
    names_sep  = "\\.",
    values_transform = list(value = as.numeric)  # mix of dbl/duration/POSIXct
  ) |> 
  pivot_wider(
    names_from  = metric,        # columns = metrics
    values_from = value          # rows = statistics
  ) |> 
  mutate(
    across(-stat, round),
    across(contains("duration"), as.duration),
    across(contains("timing"), hms::as_hms)
    )
  }, .progress = "Calculating metrics")

metrics |> 
  imap(\(x, idx) x |> 
                  gt() |> 
                  cols_label_with(fn = \(x) x |> 
                               str_replace_all("_", " ") |> 
                               str_to_sentence()) |> 
                  tab_header(idx |> 
                              str_replace_all("_", " ") |> 
                               str_to_sentence()
                             )
                  
       )

metrics |> 
  iwalk(\(x, idx) {
       write_csv(x, glue("../results/tables/LightLogR/metrics/{idx}.csv"))
       write_csv(x |> mutate(across(-stat, as.numeric)), 
                 glue("../results/tables/LightLogR/metrics/{idx}_numeric.csv"))
    },
       .progress = "Writing metrics")
```

## Timeline

### Data

Stats for each time slot from 8:00 until 18:30.

```{r}
#| label: calculate timeline stats
time_stats <-
  data |>
  map({
    \(x) x |>
  # filter(Sequence %in% c("0001", "0002", "0003", "0004")) |>
      select(-Sequence) |> 
  aggregate_Datetime(
    unit = "30 mins",
    mean = mean(MEDI, na.rm = TRUE),
    median = median(MEDI, na.rm = TRUE),
    sd = sd(MEDI, na.rm = TRUE),
    min = min(MEDI, na.rm = TRUE),
    max = max(MEDI, na.rm = TRUE),
    q25 = quantile(MEDI, 0.25, na.rm = TRUE),
    q75 = quantile(MEDI, 0.75, na.rm = TRUE),
    n. = n()
    ) |> 
    select(-MEDI) |> 
      rename(n = n.)
  }, .progress = "Calculating Timestats")

time_stats |> 
  imap(\(x, idx) x |> 
                  gt() |> 
                  cols_label_with(fn = \(x) x |> 
                               str_replace_all("_", " ") |> 
                               str_to_sentence()) |> 
                  tab_header(idx |> 
                              str_replace_all("_", " ") |> 
                               str_to_sentence()
                             )
                  
       )

time_stats |> 
  iwalk(\(x, idx) {
       write_csv(x, glue("../results/tables/LightLogR/timeline/{idx}.csv"))
    },
       .progress = "Writing timeline")

```

### Plot

```{r}
#| label: create plots
caption <- glue("<strong>mean</strong>, <strong style = 'color:{alpha('#0073C2FF', 
          alpha = 1)}'>IQR</strong>, <strong style = 'color:{alpha('#0073C2FF', 
          alpha = 0.25)}'>range</strong>")

time_stats_plot <-
  time_stats |>
  imap({
   \(x, idx) x |> 
  gg_day(mean, geom = "blank",
         facetting = FALSE,
         y.axis.label = "melanopic EDI (lx)"
         ) |> 
  gg_photoperiod(coordinates) +
  geom_ribbon(aes(ymin = min, ymax = max), fill = "#0073C2FF", alpha = 0.25) +
  geom_ribbon(aes(ymin = q25, ymax = q75), fill = "#0073C2FF") +
  geom_line() +
  scale_y_continuous(trans = "symlog", 
                     labels = expression(0,10^0,10^1, 10^2, 10^3, 10^4, 10^5),
                     breaks = c(0, 10^0:5)) +
  geom_point() +
      coord_cartesian(ylim = c(1,10^5)) +
      labs(caption = caption, title = idx |> 
                                        str_replace_all("_", " ") |> 
                                        str_to_sentence()) +
      theme_sub_plot(caption = element_textbox_simple())
  })

time_stats_plot

time_stats_plot |> 
  iwalk(\(x, idx) {
       ggsave(plot = x, glue("../results/plots/LightLogR/Timeline_{idx}.pdf"),
              height = 4, width = 6)
    },
       .progress = "Saving timeline plots")

```

