---
title: "LightLogR for metric calculation"
format: 
  html:
    self-contained: true
    code-link: true
    code-tools: true
params:
  n_columns: null
---

## Preface

This script summarizes simulation runs of light exposure in terms of light exposure metrics and time-based values, both for tables and plots. The relevant data are stored in files that conform to `results/.*_sequence.csv`. Outputs are stored in `results/plots/LightLogR` and `results/tables/LightLogR`.

## Setup

```{r}
#| label: setup
#| message: false
library(tidyverse)
library(LightLogR)
library(gt)
library(glue)
library(here)
tz <- "America/Toronto"
```

## Import

```{r}
#| label: import
path <- "../results/tables/sequences"
files <- list.files(path, pattern = "_sequence.csv", full.names = TRUE)

n_columns <- params$n_columns
if (
  !is.null(n_columns) &&
    (
      !is.numeric(n_columns) ||
        length(n_columns) != 1L ||
        is.na(n_columns) ||
        n_columns < 1 ||
        n_columns %% 1 != 0
    )
) {
  stop("`n_columns` must be NULL or a positive whole number.")
}

read_sequence_file <- \(file) {
  if (is.null(n_columns)) {
    read_csv(file)
  } else {
    read_csv(file, col_select = all_of(seq_len(n_columns)))
  }
}

data <-
  files |>
  set_names() |>
  map(
    read_sequence_file,
    .progress = "Loading in data"
  )
data <-
  data |>
  map({
    \(x) x |>
      rename(Datetime = "...1") |>
      mutate(Datetime =
               lubridate::ymd_hms(
                 glue(
                   "2023-06-16 {
                  hms::hms(minutes = seq(0, 30*21, by = 30), hours = rep(8, 22))
             }"
                 ),tz = tz
               )) |>
      pivot_longer(
        -Datetime,
        names_to = "Sequence",
        names_prefix = "seq_",
        values_to = "MEDI"
      )
  }, .progress = "Preparing loaded data")

names(data) <- names(data) |> basename() |> str_remove("_sequence\\.csv")
```

## Metrics

Metrics to calculate, all based on the calculated values for melanopic EDI (lx):

- $TAT_{250}$ (Time above 250, lx)
- $TAT_{1000}$ (Time above 1000, lx)
- $Dose$ (Dose, lx·h)
- $MLiT_{250}$ (Mean timing of light above 250, HH:MM)
- $FLiT_{250}$ (First timing of light above 250, HH:MM)

```{r}
#| label: calculate metrics
metrics_raw <-
  data |>
  map({
    \(x) x |>
  group_by(Sequence) |> 
  summarize(
    dose(MEDI, Datetime, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 250, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 1000, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, threshold = 10000, as.df = TRUE),
  )
  }, .progress = "Calculating metrics")

metrics <-
  metrics_raw |>
  map({
    \(x) x |>
  summarize(
    across(dose:duration_above_10000,
           .fns = list(mean = ~mean(.x, na.rm = TRUE),
                       median = ~median(.x, na.rm = TRUE),
                       sd = ~sd(.x, na.rm = TRUE),
                       min = ~min(.x, na.rm = TRUE),
                       max = ~max(.x, na.rm = TRUE),
                       q25 = ~quantile(.x, 0.25, na.rm = TRUE),
                       q75 = ~quantile(.x, 0.75, na.rm = TRUE)
                       ),
           .names = "{.col}.{.fn}"
           )
    ) |>
  pivot_longer(
    cols = everything(),
    names_to   = c("metric", "stat"),
    names_sep  = "\\.",
    values_transform = list(value = as.numeric)  # mix of dbl/duration/POSIXct
  ) |>
  pivot_wider(
    names_from  = metric,        # columns = metrics
    values_from = value          # rows = statistics
  ) |>
  mutate(
    across(-stat, round),
    across(contains("duration"), as.duration),
    across(contains("timing"), hms::as_hms)
    )
  }, .progress = "Calculating metrics")

metrics |> 
  imap(\(x, idx) x |> 
                  gt() |> 
                  cols_label_with(fn = \(x) x |> 
                               str_replace_all("_", " ") |> 
                               str_to_sentence()) |> 
                  tab_header(idx |> 
                              str_replace_all("_", " ") |> 
                               str_to_sentence()
                             )
                  
       )

metrics_raw |> 
  iwalk(\(x, idx) {
       write_csv(x, glue("../results/tables/LightLogR/metrics_raw/{idx}.csv"))
       write_csv(x |> mutate(across(-Sequence, as.numeric)), 
                 glue("../results/tables/LightLogR/metrics_raw/{idx}_numeric.csv"))
    },
       .progress = "Writing metrics")
```
